\large How resource abundance and stochasticity affect organisms' range sizes

### 2025-02-17 writing/appendix-B-simulations.Rmd
"resources" to "distribution of R"; decr fig size

---
title: \large How resource abundance and stochasticity affect organisms' range sizes
subtitle: "Appendix B: Simulations"
# author:
#   - name: Stefano Mezzini
#     email: stefano.mezzini@ubc.ca
#     institute: [biol, braes]
#   - name: Chris H. Fleming
#     institute: [ucf, smith]
#   - name: E. Patrícia Medici
#     institute: [tapir, iucn, escas]
#   - name: Michael J. Noonan
#     email: michael.noonan@ubc.ca
#     institute: [biol, braes, cmps]
# institute:
#   - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
#   - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
#   - ucf: Department of Biology, University of Central Florida, Orlando, Florida 32816, United States.
#   - smith: Smithsonian Conservation Biology Institute, National Zoological Park, 1500 Remount Rd., Front Royal, VA 22630, United States.
#   - tapir: Lowland Tapir Conservation Initiative (LTCI), Instituto de Pesquisas Ecológicas (IPÊ), Rodovia Dom Pedro I, km 47, Nazaré Paulista, São Paulo 12960-000, Brazil.
#   - iucn: IUCN SSC Tapir Specialist Group (TSG), Campo Grande, Brazil.
#   - escas: Escola Superior de Conservação Ambiental E Sustentabilidade (ESCAS/IPÊ), Rodovia Dom Pedro I, km 47, Nazaré Paulista, São Paulo 12960-000, Brazil.
#   - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'hr-resource-stoch.bib'
# csl: 'the-american-naturalist.csl'
fontsize: 12pt
header-includes:
    - \renewcommand{\figurename}{Figure B\!\!} # for "Figure Bx."
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font=scriptsize} # smaller caption text
    - \usepackage{hanging} # for hanging indents in references
subparagraph: true # needed for \usepackage[small]{titlesec}
output:
  bookdown::pdf_document2:
    toc: false
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua' # for author affiliations
      - '--lua-filter=author-info-blocks.lua' # for author affiliations
editor_options: 
  chunk_output_type: console
---

@@

Within a Bayesian framework, Directed Acyclical Graphs (DAGs) can be used to infer causality [@mcelreath_statistical_2016]. The DAG below can be interpreted as follows: $\e(R)$ and $\var(R)$ are the (controlled) variables that determine resources, $R$, and, consequently, an organism's home-range size, $H$. Additionally, as indicated by the arrow going from $\e(R)$ to $\var(R)$, the effect of $\var(R)$ depends of the value of $\e(R)$ (and vice-versa, but arrows in DAGs have to be unidirectional).

```{r dag, eval=TRUE, echo = FALSE, fig.cap="Directed Acyclical Graph assumed for inferring the causal effects of $\\e(R)$ and $\\var(R)$ on $H$.", fig.height=4, fig.width=6, out.width='100%'}
dagify(H ~ R,
       R ~ E + V,
       outcome = 'H',
       labels = c('H' = 'H', 'R' = 'R', 'E' = 'E(R)', 'V' = 'Var(R)'),
       coords = list(x = c(R = 0, E = -1, V = 1, H = 0),
                     y = c(R = 1, E = 2, V = 2, H = 0))) %>%
  tidy_dagitty() %>%
  mutate(col = case_when(name == 'E' ~ '1',
                         name == 'EN' ~ '1',
                         name == 'V' ~ '2',
                         name == 'VN' ~ '2',
                         name == 'H' ~ '3',
                         TRUE ~ '4')) %>%
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = col), size = 20, alpha = 0.3) +
  geom_dag_edges() +
  geom_dag_text(aes(label = label), color = 'black', size = 4.5) +
  scale_color_manual(values = c(pal[1:3], 'black')) +
  theme_dag() +
  theme(legend.position = 'none')
```

>>>>

---
title: \large How resource abundance and stochasticity affect organisms' range sizes
subtitle: "Appendix B: Simulations"
# author:
#   - name: Stefano Mezzini
#     email: stefano.mezzini@ubc.ca
#     institute: [biol, braes]
#   - name: Chris H. Fleming
#     institute: [ucf, smith]
#   - name: E. Patrícia Medici
#     institute: [tapir, iucn, escas]
#   - name: Michael J. Noonan
#     email: michael.noonan@ubc.ca
#     institute: [biol, braes, cmps]
# institute:
#   - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
#   - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
#   - ucf: Department of Biology, University of Central Florida, Orlando, Florida 32816, United States.
#   - smith: Smithsonian Conservation Biology Institute, National Zoological Park, 1500 Remount Rd., Front Royal, VA 22630, United States.
#   - tapir: Lowland Tapir Conservation Initiative (LTCI), Instituto de Pesquisas Ecológicas (IPÊ), Rodovia Dom Pedro I, km 47, Nazaré Paulista, São Paulo 12960-000, Brazil.
#   - iucn: IUCN SSC Tapir Specialist Group (TSG), Campo Grande, Brazil.
#   - escas: Escola Superior de Conservação Ambiental E Sustentabilidade (ESCAS/IPÊ), Rodovia Dom Pedro I, km 47, Nazaré Paulista, São Paulo 12960-000, Brazil.
#   - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'hr-resource-stoch.bib'
fontsize: 12pt
header-includes:
    - \renewcommand{\figurename}{Figure B\!\!} # for "Figure Bx."
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font=scriptsize} # smaller caption text
    - \usepackage{hanging} # for hanging indents in references
subparagraph: true # needed for \usepackage[small]{titlesec}
output:
  bookdown::pdf_document2:
    toc: false
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua' # for author affiliations
      - '--lua-filter=author-info-blocks.lua' # for author affiliations
editor_options: 
  chunk_output_type: console
---

@@

Within a Bayesian framework, Directed Acyclical Graphs (DAGs) can be used to infer causality [@mcelreath_statistical_2016]. The DAG below can be interpreted as follows: $\e(R)$ and $\var(R)$ are the (controlled) variables that determine the distribution of resources, $R$, which, in turn determines an organism's home-range size, $H$.

```{r dag, eval=TRUE, echo = FALSE, fig.cap="Directed Acyclical Graph assumed for inferring the causal effects of $\\e(R)$ and $\\var(R)$ on the distributions of $R$ and $H$ in the simulations.", fig.height=4, fig.width=4, out.width='50%'}
dagify(H ~ R,
       R ~ E + V,
       outcome = 'H',
       labels = c('H' = 'H', 'R' = 'R', 'E' = 'E(R)', 'V' = 'Var(R)'),
       coords = list(x = c(R = 0, E = -1, V = 1, H = 0),
                     y = c(R = 1, E = 2, V = 2, H = 0))) %>%
  tidy_dagitty() %>%
  mutate(col = case_when(name == 'E' ~ '1',
                         name == 'EN' ~ '1',
                         name == 'V' ~ '2',
                         name == 'VN' ~ '2',
                         name == 'H' ~ '3',
                         TRUE ~ '4')) %>%
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = col), size = 20, alpha = 0.3) +
  geom_dag_edges() +
  geom_dag_text(aes(label = label), color = 'black', size = 4.5) +
  scale_color_manual(values = c(pal[1:3], 'black')) +
  theme_dag() +
  theme(legend.position = 'none')
```
